On consistency scores in text data with an implementation in R

### 2020-09-02 outputs/documents/paper_ngrams.Rmd
ngram paper updated

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying words that are commonly found together. In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections. For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset.

@@

@@

Next, the `make_internal_consistency_dataset` function turns the individual tokens into tri-grams, and keeps only the tri-grams that appears in the data for more than once. The common tri-grams are preserved in a tibble:

```{r, eval=FALSE}
# Create ngrams from the tokens
toks_ngram <- 
  quanteda::tokens_ngrams(tokens_from_example, n = 3)

@@

Finally, the `make_internal_consistency_dataset` function splits the tri-grams to two parts: the previous two words and the last word. In the end, a tibble is created with each tri-gram containing its original three-word sequence, the first two words, and the last word:

```{r, eval=FALSE}
# Create a tibble that has the first two words in one column then the third
all_tokens <-
  all_tokens %>%
  dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
                first_words = stringr::word(tokens, start = 1, end = 2),
                last_word = stringr::word(tokens, -1),
                tokens = stringr::str_replace_all(tokens, " ", "_"),
                first_words = stringr::str_replace_all(first_words, " ", "_")
  ) %>%
  dplyr::rename(last_word_expected = last_word) %>%
  dplyr::select(-n)
```

@@

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function. Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tri-gram tokens, then splits the tri-grams and preserves them into a tibble:

@@

# Create ngrams from the tokens with errors
toks_ngram_with_errors <- 
  quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 3)
all_tokens_with_errors <- 
  tibble::tibble(tokens = toks_ngram_with_errors[[1]])

all_tokens_with_errors <-
  all_tokens_with_errors %>%
  dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
                first_words = stringr::word(tokens, start = 1, end = 2),
                last_word = stringr::word(tokens, -1),
                tokens = stringr::str_replace_all(tokens, " ", "_"),
                first_words = stringr::str_replace_all(first_words, " ", "_")
  )
```

@@

The `generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidate to the text errors:

@@

The first step is to turn the body of text into the internal consistency dataset. The generated internal consistency dataset is a tibble that contains the tri-grams that appear in the internal text data more than once. Since only `there_was_no` has more than one occurrence, it is the only reference tri-gram in the internal consistency dataset:

@@

## $`unexpected words`
## # A tibble: 1 x 4
##   first_words last_word last_word_expected as_expected
##   <chr>       <chr>     <chr>              <lgl>      
## 1 there_was   na        no                 FALSE

@@

To get the external consistency score, we will employ the `generate_external_consistency_score` function. The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset. The function identifies the word "na" as an unexpected word, and generates the external consistency score as zero. Because the external dataset is larger, there are several words generated as the replacement candidates of "na":

@@

## aRianna::generate_external_consistency_score(text_to_evaluate)
## $`external consistency`
## # A tibble: 1 x 3
##   as_expected     n consistency
##   <lgl>       <int>       <dbl>
## 1 FALSE           2           0

## $`unexpected words`
## # A tibble: 11 x 4
##    first_words last_word last_word_expected as_expected
##    <chr>       <chr>     <chr>              <lgl>      
##  1 there_was   na        a                  FALSE      
##  2 there_was   na        an                 FALSE      
##  3 there_was   na        no                 FALSE      
##  4 there_was   na        not                FALSE      
##  5 there_was   na        nothing            FALSE      
##  6 there_was   na        of                 FALSE      
##  7 there_was   na        one                FALSE      
##  8 there_was   na        only               FALSE      
##  9 there_was   na        plenty             FALSE      
## 10 there_was   na        someone            FALSE      
## 11 there_was   na        something          FALSE

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. As demonstrated in our application, we use a tri-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates. However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morining' or 'work' is to be suggested as the next word after 'good'.

>>>>

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying words that are commonly found together. In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections. For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset. To calculate consistency scores and generate word candidates for text corrections, we use mixed sequences of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.

@@

In our application, we use a model that has a mixture of 5-grams, 4-grams and tri-grams to provide word replacements to a text error. Longer n-grams such as 5-grams would capture more context of the sequence and potentially generate more accurate predictions. However, there is less probability for a longer sequence of words to fit a specific combination. Therefore, shorter n-grams should still be included to preserve generalizability of the predictions.

Consider the common English idiom "elephant in the room". If by mistake, the idiom is written as "elephant in the roon" and correction is needed for the last word, we can apply both a tri-gram model or a 4-gram model to predict the candidate word:

> Using tri-gram model: "in the `<prediction>`"

> Using 4-gram model: "elephant in the `<prediction>`"

In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the prediction, and the word "room" is predicted by the 4-gram model. However, there are times people would simply describe the location of an elephant, for example "elephant in the pond". This combination of this sequence as a 4-gram might not be in a comparison dataset. Therefore, we have to also rely on the tri-gram models to provide other word candidates including "in the pond".

@@

Next, the `make_internal_consistency_dataset` function turns the individual tokens into 5-grams, 4-grams and tri-grams, and keeps only the n-grams that appears in the data for more than once. The common n-grams are preserved in a tibble:

```{r, eval=FALSE}
# Create ngrams from the tokens
toks_ngram <- 
  toks_ngram <- quanteda::tokens_ngrams(tokens_from_example, n = 5:3)

@@

Finally, the `make_internal_consistency_dataset` function splits the n-grams to two parts: the previous $n-1$ words and the last word. In the end, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:

```{r, eval=FALSE}
# Create a tibble that has the first two words in one column then the third
all_tokens <-
  all_tokens %>%
  dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
                first_words = stringr::word(tokens, start = 1, end = -2),
                last_word = stringr::word(tokens, -1),
                tokens = stringr::str_replace_all(tokens, " ", "_"),
                first_words = stringr::str_replace_all(first_words, " ", "_")
  ) %>%
  dplyr::rename(last_word_expected = last_word) %>%
  dplyr::select(-n)
```

@@

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function. Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams and preserves them into a tibble:

@@

# Create ngrams from the tokens with errors
toks_ngram_with_errors <- 
  quanteda::tokens_ngrams(tokens_from_example_with_errors, n = 5:3)

all_tokens_with_errors <- 
  tibble::tibble(tokens = toks_ngram_with_errors[[1]])

all_tokens_with_errors <-
  all_tokens_with_errors %>%
  dplyr::mutate(tokens = stringr::str_replace_all(tokens, "_", " "),
                first_words = stringr::word(tokens, start = 1, end = -2),
                last_word = stringr::word(tokens, -1),
                tokens = stringr::str_replace_all(tokens, " ", "_"),
                first_words = stringr::str_replace_all(first_words, " ", "_"))
```

@@

The `generate_internal_consistency_score` function also lists the identified text errors and generates replacement candidates to the text errors:

@@

The first step is to turn the body of text into the internal consistency dataset. The generated internal consistency dataset is a tibble that contains the 5-grams, 4-grams and tri-grams that appear in the internal text data more than once. Since only `there_was_no` has more than one occurrence, it is the only reference tri-gram in the internal consistency dataset:

@@

## $`unexpected words`
## # A tibble: 1 x 3
##   first_words last_word last_word_expected 
##   <chr>       <chr>     <chr>                 
## 1 there_was   na        no

@@

To get the external consistency score, we will employ the `generate_external_consistency_score` function. The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset. Here, the text to evaluate is "there was no possibiliti", with possibility being wrongly spelled. The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score as 0.833. Because the external dataset is larger, there are several words generated as the replacement candidates of "possibiliti". The replacements based on 4-grams are displayed first followed by results based on tri-grams:

@@

## aRianna::generate_external_consistency_score(text_to_evaluate)
## $`external consistency`
## # A tibble: 1 x 3
##   as_expected     n consistency
##   <lgl>       <int>       <dbl>
## 1 FALSE           2       0.833
## 2 TRUE           10       0.833

## $`unexpected words`
## # A tibble: 11 x 3
##    first_words  last_word    last_word_expected 
##    <chr>        <chr>        <chr>                    
##  1 there_was_no possibiliti  evidence          
##  2 there_was_no possibiliti  immediate         
##  3 there_was_no possibiliti  infrastructure    
##  4 there_was_no possibiliti  one               
##  5 there_was_no possibiliti  possibility       
##  6 there_was_no possibiliti  sound             
##  7 there_was_no possibiliti  way               
##  8 was_no       possibiliti  good              
##  9 was_no       possibiliti  longer            
## 10 was_no       possibiliti  more              
## 11 was_no       possibiliti  wonder 
```

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. As demonstrated in our application, we use a tri-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates. However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morning' or 'work' is to be suggested as the next word after 'good'.
